---
title: "SEM Course - Final Project"
author: "Noa Valansi & Tomer Zipori"
date: 2024-09-15
execute: 
  warning: false
  message: false
  cache: true
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    toc-location: right
editor: visual
---

# Setup

```{r}
library(tidyverse)
library(lavaan)
library(ggfortify)
library(tidySEM)
```

# Data loading and wrangling

```{r}
data <- read.csv("../final/MentalHealthSurvey.csv")

data_clean <- data |>
  select(where(is.numeric))
```

We create two latent variables:
* (reversed) Well_being from `depression`, `anxiety` and `isolation`.
* Stress from `academic_workload` and `academic_pressure`.

# Measurement model

```{r}
meaurement_model <- '
  ## latent variable definitions (CFA)
  Stress =~ academic_workload + academic_pressure
  Well_being =~ depression + anxiety + isolation

  ## Covariances
  Well_being ~~ Stress

  ## Regressions
   
  ## errors
   
  ## autocorrelations
  
  ## self-regression (do not forget!)
'
```

## Model fitting and output

```{r}
fit_meas <- cfa(meaurement_model, data = data_clean)

summary(fit_meas, fit.measures = TRUE, standardize = F)
```

Factors constructing both latent variables show significant loadings, indicating that they contribute to the measurement of them. As can be expected, positive and significant correlation between the two latent variables.

## Goodness of fit

```{r}
fitMeasures(fit_meas, output = "text",
            fit.measures = c("nfi","nnfi","tli","cfi","gfi","rmsea"))
```

Model seems to fit the data well.

Comparing the measurement model to the saturated (just identified) model:

```{r}
anova(fit_meas)
```

Measurement model does not fit significantly better to the data than the saturated model.

Using `modificationindices()` to see what changes to the model could improve it's fit:

```{r}
modificationindices(fit_meas, sort. = TRUE, maximum.number = 5)
```

It seems that estimating the covariance between `academic_workload` and `isolation`, and between `depression` and `anxiety`, as well as adding `isolation` to the measurement of `Stress` will improve model fit.

Reliability of each latent variable measurement:

```{r}
semTools::reliability(fit_meas)
```

## Visualization

Measurement of `Well_being` is more reliable than the measurement of `Stress`. 

```{r}
lay <- get_layout(
  "academic_pressure", "academic_workload", "isolation", "anxiety", "depression",
  "Stress", NA, NA, NA, "Well_being",
  rows = 2
)
```

```{r}
graph_sem(fit_meas, angle = 90, layout = lay)
```

# Hypothesis 1 - Mediation

Hypothesis 1 - A mediation model: Stress predicts bad well being, which in turn predict lower study satisfaction.

## Hypothesized model

```{r}
hypothesized_model <- '
  ## latent variable definitions (CFA)
  Stress =~ academic_workload + academic_pressure
  Well_being =~ depression + anxiety + isolation

  ## Covariances

  ## Regressions
  Well_being ~ Stress
  study_satisfaction ~ Stress + Well_being
   
  ## effects
   
  ## autocorrelations
  
  ## self-regression (do not forget!)
'
```

```{r}
hypothesized_fit <- sem(hypothesized_model, data = data_clean)
```

```{r}
summary(hypothesized_fit)
```

Direct path - `Stress` is not significantly correlated with `study satisfaction` when controlling for `well being`.

Indirect path - `Well being` is significantly correlated with `Stress`, and it is also significantly correlated with `study satisfaction`.

### Goodness of fit

```{r}
fitMeasures(hypothesized_fit, output = "text",
            fit.measures = c("nfi","nnfi","tli","cfi","gfi","rmsea"))
```

Good model fit.

```{r}
anova(hypothesized_fit)
```

The hypothesized model is not a better fit than the saturated model.

### Visualiztion

```{r}
lay <- get_layout(
  NA, NA, "depression", "anxiety", "isolation", NA,
  NA, NA, NA, "Well_being", NA, NA,
  NA, "Stress", NA, NA, NA, "study_satisfaction",
  "academic_pressure", NA, "academic_workload", NA, NA, NA,
  rows = 4
)
```

```{r}
graph_sem(hypothesized_fit,
          edges = get_edges(hypothesized_fit, label = "est_std"),
          angle = 90,
          layout = lay)
```

Fixing the correlation between `Well_being` and `study_satisfaction` to 0, nullifying the indirect path.

## Restricted model

```{r}
restricted_hypothesized_model <- '
  ## latent variable definitions (CFA)
  Stress =~ academic_workload + academic_pressure
  Well_being =~ depression + anxiety + isolation

  ## Covariances
  Well_being ~~ 0 * study_satisfaction

  ## Regressions
  Well_being ~ Stress
  study_satisfaction ~ 0 * Well_being + Stress
   
  ## effects
   
  ## autocorrelations
  
  ## self-regression (do not forget!)
'
```

```{r}
restricted_hypothesized_fit <- sem(restricted_hypothesized_model, data = data_clean)
```

```{r}
summary(restricted_hypothesized_fit)
```

### Goodness of fit

```{r}
fitMeasures(restricted_hypothesized_fit, output = "text",
            fit.measures = c("nfi","nnfi","tli","cfi","gfi","rmsea"))
```

Comparing between the hypothesized model and the restricted model.

```{r}
anova(restricted_hypothesized_fit, hypothesized_fit)
```

### Visualiztion

The restricted model is a significantly better fit. The simpler model, without the indirect path is better. 

```{r}
graph_sem(restricted_hypothesized_fit,
          edges = get_edges(restricted_hypothesized_fit, label = "est_std"),
          angle = 90,
          layout = lay)
```

# Hypothesis 2 - Group analysis (interaction)

Do older students deal better with stress, do the correlation between `Stress` and `Well_being` differ for different ages?

Creating a categorical variable for age because `lavaan` doesn't support interaction terms for latent variables.
```{r}
data_clean <- data_clean |>
  mutate(age_cat = case_when(age < median(age) ~ "young",
                             age >= median(age) ~ "old",
                             .default = NA))
```

## Model fitting

```{r}
interaction_hypothesized_model <- '
  ## latent variable definitions (CFA)
  Stress =~ academic_workload + academic_pressure
  Well_being =~ depression + anxiety + isolation

  ## Covariances

  ## Regressions
  Well_being ~ Stress
   
  ## effects
   
  ## autocorrelations
  
  ## self-regression (do not forget!)
'
```

```{r}
interaction_hypothesized_fit <- sem(interaction_hypothesized_model, data = data_clean,
                                    group = "age_cat")
```

```{r}
summary(interaction_hypothesized_fit)
```

### Goodness of fit

```{r}
fitMeasures(interaction_hypothesized_fit, output = "text",
            fit.measures = c("nfi","nnfi","tli","cfi","gfi","rmsea"))
```

Very good fit.

Comparing to the saturated model:

```{r}
anova(interaction_hypothesized_fit)
```

### Visualiztion

The model is not significantly better than the saturated model.

```{r}
lay <- get_layout(NA, "Stress", NA, NA, "Well_being", NA,
                  "academic_pressure", NA, "academic_workload", "depression", "anxiety", "isolation",
                  rows = 2)
```


```{r}
graph_sem(interaction_hypothesized_fit, 
          edges = get_edges(interaction_hypothesized_fit, label = "est_std"),
          nodes = get_nodes(interaction_hypothesized_fit, label = "name"),
          layout = lay,
          angle = 90)
```

As hypothesized, correlation between `Stress` and `Well_being` is stronger among young students. Is it significant?

### Comparison of parameters

```{r}
interaction_hypothesized_model_comp <- '
  ## latent variable definitions (CFA)
  Stress =~ academic_workload + academic_pressure
  Well_being =~ depression + anxiety + isolation

  ## Covariances

  ## Regressions
  Well_being ~ c(Oeffect, Yeffect) * Stress
  
  ## Computed estimates
  gender_effect_diff := Oeffect - Yeffect
   
  ## effects
   
  ## autocorrelations
  
  ## self-regression (do not forget!)
'
```

```{r}
interaction_hypothesized_fit <- sem(interaction_hypothesized_model_comp, data = data_clean,
                                    group = "age_cat")
```

```{r}
summary(interaction_hypothesized_fit)
```

Estimated parameters for young students ($b=1.814$) and old students ($b=1.07$) does not differ significantly ($d=-0.745$, $p=0.397$).

